---
title: "stringr"
author: "George Zhou"
output: 
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: true
---

*Presented 6/19/2020*

## Introduction

The library stringr created to work with/manipulate strings in R. It is a wrapper of the library *stringi*, which uses ICU C library to provide fast, easy, and correct implementation functions.

stringr is a core library in the *tidyverse* world. You just need to install tidyverse to be able to enjoy stringr!  

## So what can stringr actually do? {.tabset .tabset-fade}

A lot actually. Here are some examples of functions in the stringr package. 

### str_count

[str_count](https://www.rdocumentation.org/packages/stringr/versions/1.4.0/topics/str_count) counts the number of strings that contain a match with the pattern given. 
```{r include=F}
library(tidyverse)
library(ggbeeswarm)
```


```{r}
string='I love R so so much. R is the best thing since sliced bread. Would you agree with me that R is the greatest thing that has happened to you?'
pattern='R'
str_count(string,pattern)
```

You are also able to use regex patterns.

```{r}
#How many capital words are there? 
string='One fish, Two fish, Red fish, Blue fish. Black fish, Blue fish, Old fish, New fish.'
pattern='[A-Z]'
str_count(string,pattern)
```


### str_replace_all

[str_replace_all](https://www.rdocumentation.org/packages/stringr/versions/1.4.0/topics/str_replace) Replaces all matched patterns within each string.

There is also a [str_replace](https://www.rdocumentation.org/packages/stringr/versions/1.4.0/topics/str_replace) function that replaces the first matched pattern in a string. 

```{r }
string='I am a proud Wolverine. I will forever and always be a Wolverine. Wolverines for life!'
pattern='Wolverine'
replacement='Badger'
str_replace_all(string,pattern,replacement)
#str_replace only changes the first match
str_replace(string,pattern,replacement)
```

```{r}
string1='He is pretty cool. She is pretty cool. Kanye is pretty cool.'
pattern1='[A-Z]{1}[a-z]*'
replacement1='George'
str_replace_all(string1,pattern1,replacement1)
```

### str_split_fixed

[str_split_fixed](https://www.rdocumentation.org/packages/stringr/versions/0.6.2/topics/str_split_fixed) splits a vector of strings into a matrix of substrings. 

```{r}
string=c('Sam Sung','Joe King','Horace Cope','Chandler Eere','Walter Slide','Ray Sing','Al E. Gator')
#n is number of pieces to return 
pattern=' '
str_split_fixed(string,pattern,n=2)
```

### str_detect and str_which, str_extract_all

[str_detect](https://www.rdocumentation.org/packages/stringr/versions/1.4.0/topics/str_detect) and [str_which](https://www.rdocumentation.org/packages/stringr/versions/1.4.0) perform similar functions as grepl and grep respectively. 

[str_extract_all](https://www.rdocumentation.org/packages/stringr/versions/0.6.2/topics/str_extract) will pull the texts that matched with the pattern. 

```{r}
string=c('Never ever!','Gonna get some cookies!','Give it to me!','You are awesome!','Up up and away!')
pattern='^[A-Z]{1}[a-z]*'
#is there a first word that is capitalized? 
str_detect(string,pattern)
#which strings has a capital first word?
str_which(string,pattern)
#extracts the matched word 
str_extract(string,pattern)
```

## So why stringr? {.tabset .tabset-fade}

stringr offers functions that may have already been covered in base R. So why use it? 

### Speed

One reason to use stringr over base R is the processing speed is much faster. 

In this example, we will using vector sentences, which is a vector of sentences.

```{r}
head(sentences)
```


```{r}
results=bench::mark(
  base=grep('The',sentences),
  stringr=str_which(sentences,'The'),iterations = 1000,time_unit = 's')
results
dat=data.frame(Method=c(rep('base',1000),rep('stringr',1000)),Time=c(results$time[[1]],results$time[[2]]))
ggplot(dat,aes(Method,Time))+
  geom_quasirandom(aes(color=Method))+
  scale_y_log10(name='Time (Seconds)')+
  scale_color_manual(values=c('red4','steelblue1'))+
  ggtitle("Time Distribution based on Method")+
  theme_minimal()+
  theme(legend.position = 'none')
```

*us* is microseconds, which equals .001 milliseconds (ms).

### Simplicity

Another reason to use stringr functions is the syntax is simpler in some cases, such as when piping. 
```{r,results='hide'}
new=sentences%>%str_detect('The')
#compared to 
old=sentences%>%grepl('The',.)
```

Many tasks that you would want to do in base R might need extra steps, whereas in stringr, you only need to use one function. 

```{r}
string=c("Thank goodness it is Friday! Friday is the best day in the world! Mondays are superior!")
pattern='Friday'
string1=unlist(strsplit(string,split ="!"))
sum(grepl(pattern,string1))

#same results, less code needed
str_count(string,pattern)
```

### Times where Base might be better

There are times where you may want to use base R instead. In some cases, stringr functions need helper functions to do its tasks properly. Base functions have built in arguments so there is no need for helper functions

```{r}
#Exact matched cases
string=c('Yay.Friday','Yay it is Friday!')
pattern='Yay.Friday'

grepl(pattern,string,fixed = T)

#same results, helper function needed
str_detect(string,fixed(pattern))
```

## Conclusion

* stringr is a library enclosed in tidyverse built to make string manipulation easy
* Advantages in using stringr is usability, speed, and efficiency
* Disadvantages include not being versatile and need of helper functions in some cases 
    
## Resources 

* [stringr website](https://stringr.tidyverse.org/) (includes stringr cheat sheet!)
* [regex documentation](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) 
* [R for Data Science Strings Chapter](https://r4ds.had.co.nz/strings.html)
* [stringi Documentation](https://cran.r-project.org/web/packages/stringi/index.html)